Classification and Regression Trees and Beyond

### 2020-07-01 pkg/partykit/inst/ULGcourse-2020/slides_tree.Rnw
finished revision of tree slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3028 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 937f161b6b4c726f83e9e0db39c1fcf3beaf8185

\textbf{Split point selection:} Optimize some criterion over all (binary) splits in $X_{j*}$.
\begin{itemize}
  \item Objective function.
  \item Two-sample test statistic or corresponding $p$-value.
\end{itemize}

@@

\textbf{Idea:}
\begin{itemize}
  \item Split $Y$ into most homogenous subgroups with respect to selected $X_{j*}$.
  \item Homogeneity captured by goodness-of-fit measure.
  \item \emph{Often:} Consider only binary splits.
  \item Trivial for binary $X_{j*}$.
  \item Otherwise typically exhaustive search over all binary splits.
  \item Possibly already done in split variable selection.
\end{itemize}

@@

\textbf{Pre-pruning:}
\begin{itemize}
\item[+] Does not require additional calculations as only already provided information from statistical test is used.
\item[+] Computationally less expensive as trees are prevented from getting too large.
\item[-] Performance depends on the power of the selected statistical test.
\end{itemize}

@@

\textbf{Post-pruning:}
\begin{itemize}
\item[+] Stable method to avoid overfitting, regardless of the power of the employed statistical test.
\item[+] Can employ information criteria such as AIC or BIC for model-based partitioning. 
\item[-] Computationally more expensive as very large trees are grown and additional (out-of-bag) evaluations are required.
\end{itemize}

@@

\subsection{Algorithms \& implementations}

\begin{frame}
\frametitle{CTree}

\begin{itemize} 
\item Conditional inference trees.
\item Flexible and general framework for unbiased recursive partitioning,
originally proposed for nonparametric modeling but also allowing for a model based approach via transformations of the response. 
\item Split variable selection based on conditional permutation tests.
\item Split point selection based on two-sample test statistics.
\item Employs pre-pruning.
\item Implemented in the R package partykit.
\end{itemize}

@@

\textbf{Note:} The CTree algorithm has been employed for all previous tree illustrations.

@@

\begin{frame}[fragile]
\frametitle{CTree}

Apply \fct{ctree} to preprocessed \code{ttnc} data:

<<ctree, echo=TRUE, eval=FALSE>>=
ct <- ctree(Survived ~ Gender + Age + Class, data = ttnc)
@

@@

Plot:

<<ctree-plot, echo=TRUE, eval=FALSE>>=
plot(ct)
@

@@

Add control arguments via \fct{ctree\_control}:

<<ctree_control, echo=TRUE, eval=FALSE>>=
ct <- ctree(Survived ~ Gender + Age + Class, data = ttnc,
            control = ctree_control(alpha = 0.01,
                                    minsplit = 20, 
                                    maxdepth = 3))
@
\end{frame}

\begin{frame}[fragile]
\frametitle{CTree}

\setkeys{Gin}{width=0.9\linewidth}
<<ctree_control_plot, echo=TRUE, eval=TRUE, fig=TRUE, width=10, height=5>>=
ct <- ctree(Survived ~ Gender + Age + Class, data = ttnc,
            control = ctree_control(alpha = 0.01, minsplit = 20, maxdepth = 3))
plot(ct)
@

@@

\begin{frame}[fragile]
\frametitle{CTree}

\textbf{Predictions:} For a (possibly new) set of observations the method \fct{predict} can be employed to obtain the corresponding predicted values of the response variable.

<<ctree_predict, echo=TRUE, eval=TRUE, results=verbatim>>=
nd <- data.frame(Class = "2nd", Gender = c("Female", "Male"), Age = "Child")
nd
predict(ct, newdata = nd)
@

@@

\begin{frame}
\frametitle{rpart}

\begin{itemize} 
\item Recursive PARTitioning.
\item Based on ideas of CART (Classification And Regression Trees).
\item Split variable and split point selection based on optimizing an objective function.
\item Classification trees: reduce number of misclassifications.
\item Regression trees: reduce objective function for numeric response.
\item Post-pruning can be employed.
\item Implemented in the R package rpart.
\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{rpart}

Apply \fct{rpart} to preprocessed \code{ttnc} data:

<<rpart, echo=TRUE, eval=TRUE>>=
rp <- rpart(Survived ~ Gender + Age + Class, data = ttnc)
@

@@

Standard plot:

<<rpart-plot, echo=TRUE, eval=FALSE>>=
plot(rp)
text(rp)
@

Visualization via partykit:

<<rpart-party-plot, echo=TRUE, eval=FALSE>>=
plot(as.party(rp))
@
\end{frame}

\begin{frame}[fragile]
\frametitle{rpart}

<<rpart-plot_eval, echo=TRUE, eval=FALSE>>=
plot(rp)
text(rp)
@

\vspace{-0.7cm}

\setkeys{Gin}{width=0.9\linewidth}
<<rpart-plot_eval, echo=FALSE, eval=TRUE, fig=TRUE, width=10, height = 6>>=
plot(rp)
text(rp)
@

@@

\begin{frame}[fragile]
\frametitle{rpart}

<<rpart-party-plot_eval, echo=TRUE, eval=FALSE>>=
plot(as.party(rp))
@

%\vspace{-0.4cm}

\setkeys{Gin}{width=0.9\linewidth}
<<rpart-party-plot_eval, echo=FALSE, eval=TRUE, fig=TRUE, width = 12>>=
plot(as.party(rp))
@
\end{frame}

\begin{frame}[fragile]
\frametitle{rpart}

Standard output: 
<<rpart-print, echo=TRUE, results=verbatim>>=
rp
@

@@

\begin{frame}[fragile]
\frametitle{rpart}

Output via partykit: 
<<rpart-party_print, echo=TRUE, results=verbatim>>=
as.party(rp)
@

@@

\begin{frame}[fragile]
\frametitle{rpart}

\textbf{Post-pruning:} Cost-complexity pruning of an rpart-object with complexity parameter \texttt{cp}.

%\medskip

<<rpart_pruning, echo=TRUE, eval=TRUE, results=verbatim>>=
rp_pruned <- prune(rp, cp = 0.01)
@

\setkeys{Gin}{width=0.6\linewidth}
<<rpart_pruning_plot, echo=TRUE, eval=TRUE, fig=TRUE, width = 8, height=5>>=
plot(as.party(rp_pruned))
@
% plot(rp_pruned)
% text(rp_pruned)

@@

\begin{frame}[fragile]
\frametitle{rpart}

\textbf{Post-pruning:} Cost-complexity pruning of an rpart-object with complexity parameter \texttt{cp}.

%\medskip

<<rpart_pruning2, echo=TRUE, eval=TRUE, results=verbatim>>=
rp_pruned <- prune(rp, cp = 0.02)
@

\setkeys{Gin}{width=0.6\linewidth}
<<rpart_pruning2_plot, echo=TRUE, eval=TRUE, fig=TRUE, width = 8, height=5>>=
plot(as.party(rp_pruned))
@
% plot(rp_pruned)
% text(rp_pruned)

@@

\begin{frame}[fragile]
\frametitle{rpart}

\textbf{Post-pruning:} Cost-complexity pruning of an rpart-object with complexity parameter \texttt{cp}.

%\medskip

<<rpart_pruning3, echo=TRUE, eval=TRUE, results=verbatim>>=
rp_pruned <- prune(rp, cp = 0.03)
@

\setkeys{Gin}{width=0.6\linewidth}
<<rpart_pruning3_plot, echo=TRUE, eval=TRUE, fig=TRUE, width =8, height=5>>=
plot(as.party(rp_pruned))
@
% plot(rp_pruned)
% text(rp_pruned)

@@

\begin{frame}
\frametitle{MOB}

\begin{itemize} 
\item Model-based recursive partitioning using M-estimation (ML, OLS, CRPS, \ldots).
\item Split variable selection based on asymptotic parameter instability tests.
\item Split point selection based on exhaustive search.
\item Employs pre-pruning, but can also use post-pruning based on AIC or BIC.
\item Implemented in the R package partykit.
\end{itemize}

@@

% \begin{frame}
% \frametitle{R software for tree models}
% 
% \textbf{rpart:} 
% \begin{itemize}
% \item R package for Recursive PARTitioning.
% \item Implementation based on ideas of CART (Classification And Regression Trees).
% \item Classification trees: reduce number of misclassifications.
% \item Regression trees: reduce objective function for numeric response.
% \end{itemize}
% 
% \vspace{0.4cm}
% 
% \textbf{partykit:} 
% \begin{itemize} 
% \item R package offering a toolkit for unbiased recursive partitioning. 
% \item Implementations of the tree algorithms CTree (Conditional Inference Trees) and MOB (Model-based Recursive Partitioning).
% \item Provides a very general framework with a unifying infrastructure.
% \item Can import rpart.
% \end{itemize}
% 
% \end{frame}

\begin{frame}[fragile]
\frametitle{MOB}

Is ``women and children first'' applied equally for all passenger classes?

Add a factor variable distinguishing women and children from male adults:
<<ttnc_treatment, echo=TRUE, eval=TRUE>>=
ttnc <- transform(ttnc, Treatment = factor(Gender == "Female" | Age == "Child", 
                                           levels = c(FALSE, TRUE),
                                           labels = c("Male&Adult", "Female|Child")))
@

Assess differences in the preferential treatment effect over the remaining covariates by applying \fct{glmtree} which employs the MOB algorithm for generalized linear models in the tree nodes:

<<glmtree, echo=TRUE, eval=TRUE, results=verbatim>>=
glmt <- glmtree(Survived ~ Treatment | Class + Gender + Age,
                data = ttnc, family = binomial, alpha = 0.01)
@

<<glmtree_print, echo=TRUE, eval=FALSE>>=
glmt
@
\end{frame}

\begin{frame}[fragile]
\frametitle{MOB}
\vspace{-0.4cm}
\small
<<glmtree_print_eval, echo=FALSE, eval=TRUE, results=verbatim>>=
glmt
@
\end{frame}

\begin{frame}[fragile]
\frametitle{MOB}
\setkeys{Gin}{width=0.99\textwidth}
<<glmtree-plot, echo=TRUE, eval=TRUE, fig=TRUE, width=11, height=5>>=
plot(glmt)
@
\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Further R packages based on partykit:}
\begin{itemize}
  \item evtree (evolutionary trees),
  \item model4you (personalised trees),
  \item disttree (distributional trees),
  \item stablelearner (stability assessment),
  \item \ldots
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Further R packages:}
\begin{itemize}
  \item mvpart (multivariate CART),
  \item RWeka (J4.8, M5', LMT),
  \item and many more (C50, quint, stima, \dots).
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{R software for tree models}

\textbf{Furthermore:} Tree algorithms/software without R interface, e.g.,
%
\begin{itemize}
  \item QUEST,
  \item GUIDE,
  \item LOTUS,
  \item CRUISE,
  \item \dots
\end{itemize}

@@

% \subsection*{References}

@@

%\vspace{-0.2cm}

>>>>

\textbf{Split point selection:} Optimize some criterion over all (binary) splits in $X_{j^*}$.
\begin{itemize}
  \item Objective function.
  \item Two-sample test statistic or corresponding $p$-value.
\end{itemize}

@@

\textbf{Idea:}
\begin{itemize}
  \item Split $Y$ into most homogenous subgroups with respect to selected $X_{j^*}$.
  \item Homogeneity captured by goodness-of-fit measure.
  \item \emph{Often:} Consider only binary splits.
  \item Trivial for binary $X_{j^*}$.
  \item Otherwise typically exhaustive search over all binary splits.
  \item Possibly already done in split variable selection.
\end{itemize}

@@

\textbf{Pre-pruning:}
\begin{itemize}
\item[+] Does not require additional calculations as only already provided information from statistical test is used.
\item[+] Computationally less expensive as trees are prevented from getting too large.
\item[--] Performance depends on the power of the selected statistical test.
\end{itemize}

@@

\textbf{Post-pruning:}
\begin{itemize}
\item[+] Stable method to avoid overfitting, regardless of the power of the employed statistical test.
\item[+] Can employ information criteria such as AIC or BIC for model-based partitioning. 
\item[--] Computationally more expensive as very large trees are grown and additional (out-of-bag) evaluations are required.
\end{itemize}

@@

\subsection{Conditional inference trees}

\begin{frame}
\frametitle{Conditional inference trees}

\textbf{CTree:}
\begin{itemize} 
  \item Employs general conditional inference (or permutation test) framework for association
    of $h(Y)$ vs.\ $g(X_j)$.
  \item Broadly applicable by choosing suitable transformations $h(\cdot)$ and $g(\cdot)$.
  \item \emph{Examples:} Univariate $Y$ of arbitrary scale, multivariate, model-based, etc.
  \item \emph{Default:} Pre-pruning based on direct asymptotic tests for split variable selection.
    Maximally-selected two-sample statistics for split point selection.
  \item \emph{Software:} R package \emph{partykit} (and previously: \emph{party}).
  \item \emph{Reference:} Hothorn, Hornik, Zeileis (2006).
\end{itemize}

@@

\textbf{Note:} CTree used in previous illustrations of classification and regression trees.

@@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Illustration:} Titanic survival data.

\textbf{Data:} Preprocessing from four-way contingency table \code{Titanic} in base R.

<<echo=TRUE>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
@

\textbf{CTree:} With default arguments.

<<ctree, echo=TRUE, eval=FALSE>>=
library("partykit")
ct_ttnc <- ctree(Survived ~ Gender + Age + Class, data = ttnc)
plot(ct_ttnc)
print(ct_ttnc)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, height=6, width=11>>=
plot(ct_ttnc)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

<<echo=FALSE, eval=TRUE, results=verbatim>>=
print(ct_ttnc)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Predictions:} Different types.

<<echo=TRUE, results=verbatim>>=
ndm <- data.frame(Gender = "Male", Age = "Adult", Class = c("1st", "2nd", "3rd"))
predict(ct_ttnc, newdata = ndm, type = "node")
predict(ct_ttnc, newdata = ndm, type = "response")
predict(ct_ttnc, newdata = ndm, type = "prob")
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Predictions:} Women and children first?

<<echo=TRUE, results=verbatim>>=
ndf <- data.frame(Gender = "Female", Age = "Adult", Class = c("1st", "2nd", "3rd"))
ndc <- data.frame(Gender = "Male", Age = "Child", Class = c("1st", "2nd", "3rd"))
cbind(
  Male   = predict(ct_ttnc, newdata = ndm, type = "prob")[, 2],
  Female = predict(ct_ttnc, newdata = ndf, type = "prob")[, 2],
  Child  = predict(ct_ttnc, newdata = ndc, type = "prob")[, 2]
)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Hyperparameters:} See \code{?ctree_control}.
\begin{itemize}
  \item \code{alpha}: Significance level for pre-pruning in split variable selection.
  \item \code{minsplit}: Minimum number of observations in a node for splitting.
  \item \code{minbucket}: Minimum number of observations in a terminal node.
  \item \code{maxdepth}: Maximum depth of the tree.
  \item \code{multiway}: Use multiway splits for unordered factors with $> 2$ levels?
  \item \dots
\end{itemize}

\textbf{Example:}

<<echo=TRUE, eval=TRUE>>=
ct_ttnc2 <- ctree(Survived ~ Gender + Age + Class, data = ttnc,
  alpha = 0.01, minbucket = 5, minsplit = 15, maxdepth = 4)
plot(ct_ttnc2)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\setkeys{Gin}{width=\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=13, height=6.9>>=
plot(ct_ttnc2)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Predictions:} Male children in 1st and 2nd class are now in their own terminal nodes.

<<ctree_predict, echo=TRUE, eval=TRUE, results=verbatim>>=
predict(ct_ttnc2, newdata = ndc, type = "prob")
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Evaluations:} ``As usual.''
\begin{itemize}
  \item In-sample vs. out-of-sample.
  \item Scoring rules (e.g., misclassification rates or other loss functions).
  \item Confusion matrix.
  \item Receiver operator characteristic (ROC) curve.
  \item \dots
\end{itemize}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{In-sample:} Augment learning data.

<<echo=TRUE, results=verbatim>>=
ttnc$Fit <- predict(ct_ttnc2, type = "response")
ttnc$Group <- factor(predict(ct_ttnc2, type = "node"))
@

\textbf{Confusion matrix:}

<<echo=TRUE, results=verbatim>>=
xtabs(~ Fit + Survived, data = ttnc)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Terminal nodes:} Recompute empirical survival rates.

<<echo=TRUE, results=verbatim>>=
tab <- xtabs(~ Group + Survived, data = ttnc)
prop.table(tab, 1)
@

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{More graphics:} \emph{ggplot2} support via \code{autoplot()} method from \emph{ggparty}.

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=13, height=6.9>>=
library("ggparty")
theme_set(theme_minimal())
autoplot(ct_ttnc2)
@

\subsection{Recursive partitioning}
\begin{frame}
\frametitle{Recursive partitioning}

\textbf{RPart:}
\begin{itemize} 
  \item Implements classic CART algorithm (classification and regression trees).
  \item Split variable and split point selection via exhaustive search based on objective functions.
  \item Hence biased towards split variables with many possible splits.
  \item Cross-validation-based post-pruning and no pre-pruning.
  \item \emph{Default:} Gini impurity for classification, residual sum of squares for regression.
  \item \emph{Software:} R package \emph{rpart}. Similar implementations in other languages (e.g., \emph{scikit-learn}).
  \item \emph{Reference:} Algorithm by Breiman \emph{et al.} (1984), open-source implementation by Therneau \& Atkinson (1997).
\end{itemize}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{R package:} Part of every standard R installation.

\textbf{Hyperparameters:} \code{?rpart.control}.

\textbf{RPart:} With default arguments.

<<rpart, echo=TRUE, eval=TRUE>>=
library("rpart")
rp_ttnc <- rpart(Survived ~ Gender + Age + Class, data = ttnc)
plot(rp_ttnc)
text(rp_ttnc)
print(rp_ttnc)
@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=10, height = 6>>=
plot(rp_ttnc)
text(rp_ttnc)
@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, results=verbatim>>=
print(rp_ttnc)
@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{Coercion:}
\begin{itemize}
  \item \code{as.party()} method to coerce \code{rpart} objects to \code{constparty} objects.
  \item Based on generic infrastructure for recursive partitioning in \emph{partykit}.
  \item Enables unified interface (print, plot, predict, \dots).
\end{itemize}

@@

<<echo=TRUE>>=
py_ttnc <- as.party(rp_ttnc)
plot(py_ttnc)
print(py_ttnc)
@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=10, height = 6>>=
plot(py_ttnc)
@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, results=verbatim>>=
print(py_ttnc)
@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{Post-pruning:}
\begin{itemize}
  \item Cost-complexity pruning, similar to using information criteria.
  \item Complexity parameter \code{cp} controls trade-off between reduction of objective function and
    tree size.  
  \item Carried out automatically during fitting.
  \item Additionally, ten-fold cross-validation carried out as well.
  \item Results in \code{rpart} object as \code{cptable}.
\end{itemize}

\textbf{Here:} Fitted object also has optimal cross-validation error (\code{xerror}).

@@

<<echo=TRUE, results=verbatim>>=
rp_ttnc$cptable
@

@@

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{Example:} Employ complexity parameter \code{cp = 0.1} for illustration.

<<echo=TRUE, results=verbatim>>=
prune(rp_ttnc, cp = 0.1)
@

@@

\subsection{Model-based recursive partitioning}
\begin{frame}
\frametitle{Model-based recursive partitioning}

\textbf{MOB:}
\begin{itemize}
  \item Parametric model in subgroups (maximum likelihood, least squares, \dots).
  \item Split variable selection based on asymptotic parameter instability tests.
  \item Split point selection via exhaustive search based on objective function.
  \item Significance-based pre-pruning (default) and optionally
    post-pruning based on information criteria (AIC, BIC, \dots).  
  \item \emph{Examples:} (Generalized) linear model tree, distributional trees, \dots
  \item \emph{Software:} R package \emph{partykit} (and previously: \emph{party}), various ``mobsters'' in extension packages.
  \item \emph{Hyperparameters:} \code{?mob\_control}.
  \item \emph{Reference:} Zeileis, Hothorn, Hornik (2008).
\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{Model-based recursive partitioning}

\textbf{Example:} Investigate treatment heterogeneity for ``women and children first.''

@@

<<ttnc_treatment, echo=TRUE, eval=TRUE>>=
ttnc <- transform(ttnc,
  Treatment = factor(Gender == "Female" | Age == "Child", 
    levels = c(FALSE, TRUE), labels = c("Male&Adult", "Female|Child")
  )
)
@

\textbf{Model:} Tree based on binary logit GLM for survival.

<<glmtree, echo=TRUE>>=
mob_ttnc <- glmtree(Survived ~ Treatment | Class + Gender + Age,
  data = ttnc, family = binomial, alpha = 0.01)
plot(mob_ttnc)
print(mob_ttnc)
@

@@

\begin{frame}[fragile]
\frametitle{Model-based recursive partitioning}

\setkeys{Gin}{width=0.83\textwidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=9, height=5>>=
plot(mob_ttnc)
@

@@

\begin{frame}[fragile]
\frametitle{Model-based recursive partitioning}

\vspace{-0.4cm}
\small
<<glmtree_print_eval, echo=FALSE, eval=TRUE, results=verbatim>>=
mob_ttnc
@

@@

\subsection{Evolutionary learning of globally optimal trees}
\begin{frame}
\frametitle{Evolutionary learning of globally optimal trees}

\textbf{EvTree:}
\begin{itemize}
  \item \emph{Goal:} Globally optimal partition rather than locally optimal split in each step.
  \item NP-hard optimization problem, attempt solution via evolutionary learning.
  \item Based on so-called ``fitness'' function.
  \item Essentially a penalized objective function, similar to information criteria or cost-complexity.
  \item Thus, no additional pruning step.
  \item \emph{Software:} R package \emph{evtree} (leverages \emph{partykit}).
  \item \emph{Hyperparameters:} \code{?evtree\_control}.
  \item \emph{Reference:} Grubinger, Zeileis, Pfeiffer (2014).
\end{itemize}

@@

\begin{frame}[fragile]
\frametitle{Evolutionary learning of globally optimal trees}

\textbf{EvTree:} With default parameters.

<<echo=TRUE, eval=FALSE>>=
library("evtree")
set.seed(1)
ev_ttnc <- evtree(Survived ~ Gender + Age + class, data = ttnc)
plot(ev_ttnc)
print(ev_ttnc)
@

<<echo=FALSE, eval=TRUE>>=
library("evtree")
if(file.exists("ev_ttnc.rds")) {
  ev_ttnc <- readRDS("ev_ttnc.rds")
} else {
  set.seed(1)
  ev_ttnc <- evtree(Survived ~ Gender + Age + Class, data = ttnc)
  saveRDS(ev_ttnc, file = "ev_ttnc.rds")
}
@

@@

\begin{frame}[fragile]
\frametitle{Evolutionary learning of globally optimal trees}

\setkeys{Gin}{width=0.72\textwidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=8, height=5>>=
plot(ev_ttnc)
@

@@

\begin{frame}[fragile]
\frametitle{Evolutionary learning of globally optimal trees}

<<echo=FALSE, eval=TRUE, results=verbatim>>=
ev_ttnc
@

@@

\subsection{Other algorithms}
\begin{frame}
\frametitle{Other algorithms}

\textbf{Furthermore:} Based on \emph{partykit}.
\begin{itemize}
  \item \emph{model4you:} Model-based trees/forests with personalised treatment effects.
  \item \emph{disttree:} Distributional modeling with regression trees and random forests.
  \item \emph{circtree:} Circular distributional trees and forests.
  \item \emph{glmertree:} Generalized linear mixed-effects model trees.
  \item \emph{psychotree:} Psychometric model trees (Rasch, Bradley-Terry, \dots).
  \item \emph{stablelearner:} Stability assessment of trees (and other learners).
  \item \ldots
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Other algorithms}

\textbf{Other R packages:}
\begin{itemize}
  \item \emph{RWeka:} Interface to \emph{Weka} machine learning library, provided open-source
    implementations of C4.5 (J.48) and M5 (M5'), and LMT.
  \item \emph{C50:} C5.0 (successor to C4.5) decision trees and rule-based models.
  \item Many more: \emph{mvpart}, \emph{quint}, \emph{stima}, \dots.
\end{itemize}

@@

\begin{frame}
\frametitle{Other algorithms}

\textbf{Beyond R:}
\begin{itemize}
  \item Unbiased recursive partitioning software (standalone) by Loh and co-workers: \emph{QUEST}, \emph{GUIDE}, \dots
  \item In Python: \emph{scikit-learn} implements classical machine learning algorithms,
    very limited availability of statistical inference-based algorithms.
  \item \dots
\end{itemize}

\subsection{References}

@@
